---
title: "Midterm Project"
author: "Justin Hsie"
date: 2018-10-25
output: github_document
---

####Setup

```{r setup, message = FALSE }
library(tidyverse)
library(viridis)
```

###Data Cleaning

```{r clean_data, message = FALSE}
accel_data = read_csv(file = "data/p8105_mtp_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(day = as.factor(day)) %>% 
  mutate(day = factor(day, levels = 
                        c("Sunday", "Monday", "Tuesday", 
                          "Wednesday", "Thursday", 
                          "Friday", "Saturday"), ordered = TRUE)) %>% 
  arrange(day) %>% 
  arrange(week) %>% 
  mutate(total_act = rowSums(.[,3:1440], na.rm = TRUE)) %>% 
  mutate(avg_act = rowMeans(.[,3:1440], na.rm = TRUE))
```
This code cleans the data and creates total activity and mean activity variables.
The dataset has `r nrow(accel_data)` rows and `r ncol(accel_data) - 2` variables total. It consists of week, day, and activity variables, where each activity column represents a summary of the voltage signal in a minute of the day. Some exploratory analysis included visually looking for outliers, but the outliers obvserved did not seem to affect the data much. 

###Total Activity Over Time

```{r total_act_over_time, message = FALSE}
year_plot = 
  ggplot(accel_data, aes(x = week, y = total_act, color = week)) +
  geom_point(alpha = .5) +
  geom_smooth() +
  labs(
    title = "Total Activity Over Time",
    x = "Week",
    y = "Total Activity",
    caption = "Data from Advanced Cardiac Care Center of CUMC"
  ) +
  viridis::scale_color_viridis(
    name = "Week"
  ) + 
  theme_bw() + 
  theme(legend.position = "")
year_plot
```

Using the calculated total activity variable, this plot shows total activity per day over the course of the year as well as a line that shows the overall trend. The data seems to support the hypothesis that the participant became more active over the course of the year, as the line has a positive slope that shows an upward trend in activity.

###Average Activity During Week Over Time

```{r all_weeks_avg_per_day}
week_plot = 
  ggplot(accel_data, aes(x = day, y = avg_act)) + 
  geom_boxplot(color = "orange", fill = "blue", alpha = .5) +
  labs(
    title = "Average Activity During Week Over Time",
    x = "Day",
    y = "Average Activity",
    caption = "Data from Advanced Cardiac Care Center of CUMC"
  ) +
  theme_bw()
week_plot
```

Here we can see the average activity each day of the week over the course of the year. It seems that the day of the week does affect his activity, with the highest and lowest medians being on Friday and Tuesday, respectively. Overall, there seems to be more activity from Thursday through Sunday, and less activity from Monday to Wednesday. In addition, the interquartile range seems to be the largest on Sunday, suggesting a larger variance in the participant's activity. 

###24 Hour Activity Throughout Week

```{r 24_hour_week, message = FALSE}
week_24hr = accel_data %>% 
  gather(key = daily_act, value = activity, activity_1:activity_1440) %>%
  filter(week == 30) %>% 
  mutate(daily_act = str_replace(daily_act, "activity_", "")) %>% 
  mutate(daily_act = as.numeric(daily_act) / 60)

week_24hr_plot = 
  ggplot(week_24hr, aes(x = daily_act, y = activity)) +
  geom_point(color = "orchid", alpha = .2) +
  scale_x_continuous(breaks = seq(0, 24, by = 6)) +
  facet_grid(. ~ day) +
  labs(
    title = "24 Hour Activity Week 30",
    x = "Hours",
    y = "Activity",
    caption = "Data from Advanced Cardiac Care Center of CUMC"
  ) +
  theme_bw() 
week_24hr_plot
```

Because the device records data continuously, we can see how the participant's activity changes throughout each day and clearly see when he is active or not. This plot shows the activity of the participant over 24 hours for each day during week 30 (chosen arbitrarily). For example, we can see that on Wednesday the participant is active between 6am and 11 pm. In addition, there is a steep drop in activity on Thursday suggesting the participant took the device off mid-day and later put it back on. 
Only one plot displaying one week was chosen for this condensed report for the sake of brevity, as opposed to including plots for every week of the study. In addition, condensing all the 24 hour profile data into a single plot would be too cluttered and not convey any particularly useful information. The following plot will show total activity over the year using 24 hour profiles with more clarity.

###24 Hour Activity Over Time

```{r avg_24_hour, message = FALSE}
avg_act = accel_data %>% 
  gather(key = daily_act, value = activity, activity_1:activity_1440) %>%
  mutate(daily_act = str_replace(daily_act, "activity_", "")) %>% 
  mutate(daily_act = as.numeric(daily_act) / 60)

avg_plot = 
  ggplot(avg_act, aes(x = daily_act, y = activity, color = day)) +
  facet_grid(. ~ day) +
  scale_x_continuous(breaks = seq(0, 24, by = 6)) +
  geom_smooth() +
  labs(
    title = "24 Hour Activity Over Time",
    x = "Hours",
    y = "Activity",
    caption = "Data from Advanced Cardiac Care Center of CUMC"
  ) +
  theme_bw() +
  theme(legend.position = "")
avg_plot
```

This plot takes the data from 24 hour profiles of every day for the whole year and displays a trend for each day of the week. Here we can see that on most days the participant is active between 6am and 11pm. There also appears to be consistent dips in activity in the middle of the day throughout the week, which could possibly be explained by the participant taking a nap or something of that nature. Overall, the day of the week seems play a role in overall activity, as there seems be more activity between Thursday through Sunday and less activity Monday through Wednesday. 




